### 2023-09-01 sections/2-methods.tex
complexity

@@

\subsubsection{Complexity of $k$-NN Search}
\label{paragraph:methods:knn-complexity}

To bound the complexity of $k$-NN search, we group all of the "Sieve" algorithms (Sieve, 
Sieve with Separate Centers, and Greedy Sieve) together, and handle $k$-NN by Repeated $\rho$-NN separately.

We start by considering the complexity of $k$-NN by Repeated $\rho$-NN, as this  
is a natural extension of the complexity bounds for $\rho$-NN search described in 
\cite{ishaq2019clustered}. For this algorithm, we adopt the terminology used in 
\cite{ishaq2019clustered} and \cite{yu2015entropy} and address \emph{coarse search} and \emph{fine search} separately. 
Coarse search refers to the process of identifying clusters
which have overlap with the query ball (i.e., which are likely to contain one of the $k$ nearest neighbors). 
Fine search refers to the process
of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.

@@

% Our complexity bounds depend on the same technical assumption about the self-similarity of the dataset made in 
% \cite{yu2015entropy}: if call the number of points within a given 
% radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
% are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
% we assume that for any point $p$, query $q$, and radius 
% $\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}

% For tree search, in the worst case, we must consider [lift ideas here from chess for this justification and ask 
% if this still holds for knn].

@@

% It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
% points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
% average local fractal dimension over all regions of the dataset is not too different from the local fractal dimension around any particular point.
% We rely on this fact in our estimation of $r'$.

% We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
% $d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
% and the cardinality of the ball centered at the root clustIn this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
% er's center with half its radius). While, ordinarily we compute local fractal dimension 
% by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
% a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.

% If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
% \begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}

% Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation 4 for $r'$ to get
% \begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}

>>>>

\subsubsection{Complexity of $k$-NN by Repeated $\rho$-NN Search}
\label{paragraph:methods:repeated-rnn-complexity}

Our complexity bounds for $k$-NN by Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution
(i.e. arises from the same generating phenomenon) as the rest of the data. We also rely on the same technical assumption about the self-similarity 
of the dataset made in \cite{yu2015entropy}: if call the number of points within a given 
radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
we assume that for any point $p$, query $q$, and radius 
$\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}

We find it useful to adopt the terminology used in 
\cite{ishaq2019clustered} and \cite{yu2015entropy} and address \emph{coarse search} and \emph{fine search} separately. 
Coarse search refers to the process of identifying clusters
which have overlap with the query ball (i.e., which are likely to contain one of the $k$ nearest neighbors). 
Fine search refers to the process
of identifying the $k$ nearest neighbors among the points in those clusters identified by coarse search.

In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search is $\mathcal{O}(\log{2}m)$, 
where $m$ is the metric entropy of the dataset. 
To adjust this bound for $k$-NN by Repeated $\rho$-NN, we must estimate the number of iterations 
of $\rho$-NN needed to find at least $k$ neighbors.

Because we assume that the query point is sampled from 
the same distribution as the rest of the data, we can assume that the local fractal dimension around the query
is not significantly different from the mean local fractal dimension of the dataset. If the local fractal dimension 
were perfectly uniform throughout the entire dataset, then 
the use of the factor from Equation \ref{2} would guarantee that it only takes 
one iteration of $\rho$-NN to find $k$ points. In practice, [is this the 
way I want to be going with this??]

To determine the asymptotic complexity of fine search, we must estimate $|F|$, the 
number of clusters for which we must look at all of the cluster's points. With $\rho$-NN 
search, $F$ is the union of clusters which are within a distance of $r + r_c$ from the query,
where $r$ is the search radius and $r_c$ is the radius of leaf clusters,
as in \cite{yu2015entropy}. With $k$-NN by Repeated $\rho$-NN, however, 
the value of $r$ is dependent on the distance to the $k$th nearest neighbor, which 
is query-dependent. Thus, for $k$-NN by Repeated $\rho$-NN, we adjust the definition of 
$F$ to be the union of clusters which are within a distance of $r' + r_c$ from the query, 
where $r'$ is the distance to the $k$th nearest neighbor and $r_c$ is defined as before.

~\cite{yu2015entropy} showed that $|F| \leq \gamma |B(q, r')|\left(\frac{r'+ 2r_c}{r'}\right)^d$. By definition of 
$r'$, we have that $|B(q, r')| = k$. Thus, we have that $|F| \leq \gamma k\left(\frac{r'+ 2r_c}{r'}\right)^d$.

It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
average local fractal dimension over all regions of the dataset does not differ significantly from the local fractal dimension around any particular point.
We rely on this fact in our estimation of $r'$.

We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
$d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
and the cardinality of the ball centered at the root cluster). In this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
er's center with half its radius. While, ordinarily we compute local fractal dimension 
by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.

If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
\begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}

Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation ~\ref{4} for $r'$ to get
\begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}

@@

\subsubsection{Complexity of Sieve, Sieve with Separate Centers, and Greedy Sieve}
\label{paragraph:methods:sieve-complexity}

@@

@@
